ECG Disease Classification Using Graph Convolution Networks

### 2017-11-22 Work_report/Final_presentation/btp_presentation.tex
More or less done

\begin{frame}
  \frametitle{Heart diseases considered}
  The following heart diseases are considered:
  \begin{itemize}
  \item Myocardial Infarction
  \item Cardiomyopathy \ Heart failure
  \item Bundle branch block
  \item Dysrhythmia
  \item Myocarditis
  \end{itemize}
\end{frame}

\section{CNN on ECG}
\begin{frame}
  \frametitle{Dataset}
  The dataset used for experiments is the PTB databas from Physionet repository.
  \begin{itemize}
  \item Consists of 15 channels (12 for conventional ecgs, 3 frank leads).
  \item Resolution is 16bit and sampling rate is 1kHz.
  \item Total 549 recordings from 290 subjects. 148 are diagnosed with Myocardial Infarction (MI).
  \item We use parts of each classification for training and rest for testing.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{CNN on ECG}
  A simple Convolutional Network is tested on ECG.
  \begin{itemize}
  \item Architecture adopted is 1D version of the LeNet architecture. The network is :
    $$conv -> max pool -> conv -> max pool -> fc -> fc -> softmax$$
  \item Base implementation results in accuracy around 80-87\%.
  \item On top of each convolution layer, we add a batch normalization.
  \item This gives accuracy boost upto 93\%-94\%
  \item State of the Art to the best of knowledge is 96\%
  \item It is also noted that using more number of channels doesn't lead to increase in the accuracy.
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Graph CNN (Optional)}
  Two main approaches to Graph CNN
  \begin{itemize}
  \item Spectral Approach [ChebNet, NIPS2016]
  \item Spatial Approach [MoNet, CVPR2017]
  \end{itemize}
  A limitation of Spectral Approach is that the underlying graph should be fixed, which is precisely the case for ECG.
\end{frame}
% \subsection{How to extend convolution to graphs?}
% \begin{frame}
%   \frametitle{Extending Convolutional to Graphs}
%   There are two main approaches
%   \begin{itemize}
%   \item <1->Spatial Approach :\\
%     Generalization of CNN in the spatial domain itself.
%     \begin{itemize}
%     \item <1-> Learning Convolutional Neural Networks for Graphs [ICML 2016].% \cite{DBLP:journals/corr/NiepertAK16}
%     \end{itemize}
%   \item <2-> Spectral Approach :\\
%     Using the frequency characterization of CNN and using that to generalize to Graphical domain
%     \begin{itemize}
%     \item <2->Spectral Networks and Deep Locally Connected Networks on Graphs [Bruna et al. ICLR 2014]. % \cite{DBLP:journals/corr/BrunaZSL13} %
%     \item <2->Convolutional Neural Networks on Graphs with Fast Localized Spectral Filtering [Defferrard et al. NIPS 2016] (will be the main focus) % \cite{DBLP:journals/corr/DefferrardBV16}
%     \item <2->Semi-Supervised Classification with Graph Convolutional Networks [Kipf et al. ICLR 2017] % \cite{DBLP:journals/corr/KipfW16}
%     \end{itemize}
%   \end{itemize}
% \end{frame}

>>>>

\begin{frame}
  \frametitle{Motivation}
  We wish to classify the disease of person based on raw ECG data.
  \begin{itemize}
  \item The main motivation comes from being able to capture real time ECG data and use that monitor the occurence of any heart related problem in the patient.
  \item This can also serve as a second opinion to the doctor.
  \end{itemize}
  Problems with Existing Methods:
  \begin{itemize}
  \item Signal Processing Techniques often may be computationally expensive and cannot work in real time.
  \item Need extremely high accuracy perhaps so as to not to trigger false alarms.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Heart diseases considered}
  The following heart diseases are considered:
  \begin{itemize}
  \item Myocardial Infarction
  \item Cardiomyopathy \ Heart failure
  \item Bundle branch block
  \item Dysrhythmia
  \item Myocarditis
  \end{itemize}
  Currently, we are bothered only with Myocardial Infarction which is major part of the PTB dataset (described next).
\end{frame}

\begin{frame}
  \frametitle{Methodology}
  The following methodology is followed:
  \begin{itemize}
  \item First we analyze using simple Convolutional Neural Networks [CNN] for our problem. CNN are a good starting point as they are well known for their spatial invariance which in 1D can be considered as the temporal invariance which is something required in a continuous ECG data.
  \item Next we hope to understand what things CNN is able to learn well and what it is not. In particular we want to identify some Ad-Hocs which lead to better convergence of the Neural Network.
  \item Then we want to know the bottleneck of the learning and what we can do to improve it, which introduces us to Graph CNN.
  \end{itemize}
\end{frame}

\section{CNN on ECG}
\begin{frame}
  \frametitle{Dataset}
  The dataset used for experiments is the PTB database from Physionet repository.
  \begin{itemize}
  \item Consists of 15 channels (12 for conventional ecgs, 3 frank leads).
  \item Resolution is 16bit and sampling rate is 1kHz.
  \item Total 549 recordings from 290 subjects. 148 are diagnosed with Myocardial Infarction (MI).
  \item We use parts of each classification for training and rest for testing.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{CNN on ECG}
  A simple Convolutional Network is tested on ECG.
  \begin{itemize}
  \item Architecture adopted is 1D version of the LeNet architecture. The network is :
    $$conv -> max pool -> conv -> max pool -> fc -> fc -> softmax$$
  \item Base implementation results in accuracy around 80-87\%.
  \item Some Ad-Hocs are required to get better results.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{CNN on ECG (contd.)}
  \begin{itemize}
  \item Since data is not so much, each ECG data is pre-processed by taking the samples upto fixed period only (eg. 38000 points are broken into 27 partitions).
  \item On top of each convolution layer, we add a batch normalization. This gives accuracy boost upto 93\%-94\%
  \item State of the Art to the best of my knowledge on the PTB dataset is 96\%.
  \item It is also noted that using more number of channels doesn't lead to increase in the accuracy.
  \end{itemize}
  We would like to do better.
\end{frame}

\begin{frame}
  \frametitle{Results on using CNNs on ECG}
  \begin{figure}[H]
    \centering
    \includegraphics[scale=0.5]{images/bn_before_relu_acc_94_3}
    \caption{Accuracy of validation set vs Number of epochs}
    \label{fig:1}
  \end{figure}
\end{frame}

@@

\begin{frame}
  \frametitle{Graph CNN (Optional)}
  Two main approaches to Graph CNN
  \begin{itemize}
  \item Spectral Approach [ChebNet, NIPS2016]
    \begin{itemize}
    \item Basic Idea : Use graph laplacian and do convolution operation in the spectral domain.
    \end{itemize}
  \item Spatial Approach [MoNet, CVPR2017]
    \begin{itemize}
    \item Basic Idea : Use a local manifold and learn in polar coordinates.
    \end{itemize}
  \end{itemize}
  A limitation of Spectral Approach is that the underlying graph should be fixed, which is fortunately the case for ECG.
\end{frame}

\begin{frame}
  \frametitle{ChebNet (Optional)}
  Main Ideas in ChebNet:
  \begin{itemize}
  \item Uses graph laplacian and computes convolution in the fourier domain (using graph fourier transform).
  \item Restricts convolution to a local domain (k-hops) even when using the spectral approach.
  \item Efficient computation of the graph convoluion using the Chebyshev Polynomials.
  \item Efficient coarsening and pooling operation on graphs.
  \end{itemize}
  Main Limitation:
  \begin{itemize}
  \item The graph should be fixed. Cannot learn on multiple graphs.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{ChebNet on ECG data}
  This is still Work In Progress. Code is ready and small bug fixes are to be made.
  % \begin{itemize}
  % \item Main limitation is that the code is
  % \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Main Limitations of the Approach}
  \begin{itemize}
  \item Neural Networks are in general data hungry. We would need large amounts of data. Unfortunately, annotated Medical data is not very easily available.
  \item Data is not directly interpretable and needs the eye of a professional. Makes debugging difficult.
  \item In usual applications of graph convolution networks, the number of nodes can be quite high which allows for efficiently using pooling and convolutions. For a small network like ECG with only 12 nodes it remains to be seen how much improvement the Graph CNN will give over vanilla CNN.
  \end{itemize}
\end{frame}

\section{Future Work}
\begin{frame}
  \frametitle{Future Work}
  \begin{itemize}
  \item Get the results on ECG data using ChebNet.
  \item Extend the results to larger datasets and even Holter ECG data.
  \item Extend the work on ECG data to EEG data as the problem space is extremely similar except for the added number of channels.
  \end{itemize}
\end{frame}

% \subsection{How to extend convolution to graphs?}
% \begin{frame}
%   \frametitle{Extending Convolutional to Graphs}
%   There are two main approaches
%   \begin{itemize}
%   \item <1->Spatial Approach :\\
%     Generalization of CNN in the spatial domain itself.
%     \begin{itemize}
%     \item <1-> Learning Convolutional Neural Networks for Graphs [ICML 2016].% \cite{DBLP:journals/corr/NiepertAK16}
%     \end{itemize}
%   \item <2-> Spectral Approach :\\
%     Using the frequency characterization of CNN and using that to generalize to Graphical domain
%     \begin{itemize}
%     \item <2->Spectral Networks and Deep Locally Connected Networks on Graphs [Bruna et al. ICLR 2014]. % \cite{DBLP:journals/corr/BrunaZSL13} %
%     \item <2->Convolutional Neural Networks on Graphs with Fast Localized Spectral Filtering [Defferrard et al. NIPS 2016] (will be the main focus) % \cite{DBLP:journals/corr/DefferrardBV16}
%     \item <2->Semi-Supervised Classification with Graph Convolutional Networks [Kipf et al. ICLR 2017] % \cite{DBLP:journals/corr/KipfW16}
%     \end{itemize}
%   \end{itemize}
% \end{frame}
